Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-03-18 SIR_manuscript.tex
Commenting, editing and thinking too much about algebra

The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics. In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are more likely to self-isolate. We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

Because individuals who have been tested or received a positive test are likely to self-isolate---reducing or eliminating  their potentially infectious contacts---the dynamics of an epidemic will depend on who gets tested and
why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
the more strongly it is focused on people who are infectious.
This level of focus depends in turn on the purpose and design of testing programs. 
When testing is done for the purposes of disease surveillance \citep{foddai2020base}
we might expect tests to be assigned randomly across the population, thus we characterize it as \emph{random} testing strategy. In an ideal world, some bias in selection for testing is probably inevitable, although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.

@@

When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the testing weights for people with vs.\ without the infections will depend on the relative magnitudes of the probability of having the infection-like symptoms when infected. Specifically, for \covid infection, the testing weights will depend on the relative proportions of presymptomatic or asymptomatic vs.\ symptomatic infections and the probability of having \covid-like symptoms when \emph{not} infected (e.g., due to other respiratory illnesses). Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.

@@

In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate. We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W}.
\end{equation}
Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.  Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.  Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).

>>>>

The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics. In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may do so also (possibly to a lesser extent). We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

If testing influences behavior, then epidemic dynamics will depend on patterns of who gets tested.
The impacts of testing will depend on intensity (tests performed per day), and on how strongly testing is focused on people who are infectious.
This level of focus depends in turn on the purpose and design of testing programs. 
When testing is done for the purposes of disease surveillance \citep{foddai2020base}
tests should be assigned randomly across the population, possibly with a stratified design for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.

@@

When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms. For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-sympomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid. Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.

@@

In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate. We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W}.
\end{equation}
Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.  Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.  Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).
